Foundations of community ecology: Supporting Information 1

### 2025-06-18 response_to_reviewers.md

General response to reviewer comments
================================================

@@

We contend that this kind of theory is important in ecology and evolution, even if it is not of widespread interest or appreciation (Caswell 1988). For example, even if the Price equation served no other purpose than to provide a formal, abstract, link between the fundamental equations of population and quantitative genetics (Queller 2017), it would still be a valid contribution to evolutionary biology, useful for providing conceptual clarity to theoreticians and demonstrating the logical consistency of the field. Hence, while we are optimistic about where our starting point might lead for developing new theory, we do not see this as being the most important or interesting outcome of this work.

Our equation is general and realistic. It is also simple; it couldn't be anything else given what we set out to achieve (and while it is perhaps trivial to verify in hindsight, it was not trivial to find). A fundamental equation of eco-evolutionary change must, by definition, be able to recover the fundamental equations of evolutionary biology (the Price equation) and population ecology (N_{t+1} = N_{t} + Births - Deaths). Our equation thereby defines what eco-evolutionary change is in its broadest sense; verbally, it is an interaction between individual fitness components (birth, death, and identity) and an individual characteristic.

Line numbers below refer to the new clean draft of our manuscript.

@@

Reviewer 1 comment 1 
==================================

The literature that is cited is VERY selective. There is a huge body of literature on linking ecological and evolutionary theory. High-dimensional matrix models have now been routinely used to link ecological and evolutionary dynamics by Barfield, Boots, Childs, Coulson, Ellner, Lion, Rees, Tuljapurkar and others. They have shown how by writing equations to dynamically track distributions of individual-level phenotypic traits and genotypes you can retrieve: the Price equation (including for structured populations) selection differentials and gradients, heritabilities, genetic covariances, the Euler-Lotka equation, equations for population growth, rates of evolutionary change, evolutionary end points, species interactions, and coexistence mechanisms. This literature needs citing and appropriately summarising. (It turns out it is straightforward to retrieve their equation from this modelling framework. if the authors did this, it will help reveal its limitations. One example is their delta zi terms get split into ontogenetic development among surviving individuals and genetic and non-genetic inheritance between parents and their offspring in age-structured populations).

Author response to Reviewer 1 comment 1
=====================================

We have completely rewritten our introduction (retaining only the first sentence) to address this comment and better communicate the goals of our work.

@@

Reviewer 1 is not correct in stating that these models retrieve the Price equation; the Price equation is not a predictive model but a mathematical identity that formally defines evolutionary change. These models don't retrieve the Price equation so much as the Price equation is necessarily embedded within them. The Price equation is not a simplified version of, e.g., Coulson and Tuljapurkar (2008), which includes a version of the Price equation that takes into account stage structure -- the latter is a specific case that is derived from the more abstract and exact former by making useful simplifying assumptions to generate novel predictions. The Price equation is also not just a generic mathematical function incorporating relevant evolutionary variables (as in, e.g., Lion and Patel examples mentioned above). Price (1970) is derived from first principles in evolutionary theory to define evolution in the same way that N_{t+1} = N_{t} + Births - Deaths defines ecological change.

Similarly, we attempt to define what is most fundamental to eco-evolutionary change. Our equation 1 does this formally, defining eco-evolutionary change in its most abstract and exact form as an interaction between fitness across different demographic components and individual characteristics. From eqn 1, the exact equations defining population and evolutionary change can be recovered. The simplicity of our equation is deceptive, and this is underscored by the fact that despite having never been discovered (to our knowledge), the equation can be used to derive the two most fundamental equations in ecology and evolution. As noted by Frank (1995, p. 381) in reference to Price's work, “[i]t is always difficult, in retrospect, to see the originality and insight of a simple idea”. While we certainly do not claim our work here to be field-changing, or as versatile as modelling tools from the abovementioned papers, we do believe that it is original and insightful in defining eco-evolutionary change.

In summary, are familiar with the literature pointed out by Reviewer 1, and we would even (quite happily) argue that this literature includes models that are highly general tools for modelling complex eco-evolutionary dynamics. The whole point of introducing the paper the way that we did was to illustrate that we are trying to ask a different question than the ones in those papers. We obviously need to try a different approach, and we have done this by completely rewriting the introduction to address the points made above.

@@

Reviewer 1 comment 2
==================================

The equation that is developed is not dynamically sufficient. This means that without modification it is fairly limited in describing what happens over an individual time step. This is no surprise, because the Price is simply a tautology that is also not dynamically sufficient. The authors may have derived an equation that is worthy of publication. They need to be honest on its limitations (also see point 1).

Author response to Reviewer 1 comment 2
==================================

We now state this more clearly in the text (L245-246).

Reviewer 1 is of course correct that our eqn 1 is not dynamically sufficient. It couldn't possibly be, for exactly the reason that Reviewer 1 points out. The Price equation is a tool for unifying evolutionary models and providing conceptual clarity, which is necessarily achieved at the cost of any and all predictive power for a given system. Models need to make (simplifying, specific system) assumptions to be predictive (Levins 1966, Frank 2012), and a fundamental equation, by its nature, needs to be free of such assumptions and focus only on what is essential. The relevance of our equation rests entirely on its power to unify fundamental equations of ecology and evolution and provide conceptual clarity, and its limitations are the same as any other fundamental equation.

@@

The authors massively overstate the importance of their equation, starting with the title of "Darwin's dream". There is no evidence I am aware of that Darwin ever thought much about equations. The authors then talk about scientific breakthroughs, attempting to place their equation with other great equations of science. Astonishing self-belief, and not warranted given how much has been done in this space already and the limitations of their formulae. The paper needs toning down significantly. The equation is mildly interesting, but it is not field-changing. This self belief may come from the authors having been working in an echo chamber. This is even hinted at in the acknowledgements and author contributions.

If the authors were to say they have found a way to write an equation that links the Price equation to the equation that describes change in population size from one time step to the next, that would be an honest and critical placement of their work in the broader literature. Explaining why it might be useful would help too.

Author response to Reviewer 1 comment 3
==================================

We have changed the title and rewritten the introduction. Our goal was to explain the importance of conceptual unification in the sciences (which, in our experience, is often underappreciated in the life sciences) without overselling our present results. In our revised introduction and discussion, we have toned down our own contribution while also not losing our central point that this kind of theory is important in science (i.e., *useful*, but because it demonstrates logical consistency and provides conceptual clarity -- not just because you can use it in the same way as a predictive model).

@@

What we have shown is a logical consistency between the fundamental equations of ecology and evolution (and, as far as we can tell, this is not disputed by either reviewer). This logical consistency has been assumed by theoreticians, and perhaps quite reasonably. But as a discipline, it is still important to do our due diligence and show that the theoretical foundations of ecology and evolution are solid. For example, within evolutionary biology, we can model evolutionary change using a population genetics framework or a quantitative genetics framework. In a population genetics framework, the average excess equation describes selection in terms of changes in allele frequencies, whereas in the quantitative genetics framework, the Breeder's equation describes selection in terms of changing quantitative traits (Queller 2017). These two equations are very important for modelling evolutionary change and predicting the consequences of selection. But an important role of theory is to demonstrate how (and why) these two equations are logically consistent, i.e., that they follow without contradiction from the same starting point that formally defines evolutionary change in any system. The Price equation and the framework that it establishes (Gardner 2020, Baravalle et al. 2025) is that starting point; it defines what evolution is in the broadest sense, meaning that any specific model of evolutionary change must be possible to derive from it. If a model of evolution cannot be put into the framework of the Price equation -- at least in principle -- then it is fundamentally flawed.

Hence, while the Price equation has very little direct relevance to the work of most evolutionary biologists, it is fundamental in the sense that it binds together all of evolutionary theory by defining what evolutionary change is (and is not). Our equation 1 fulfils the same criteria: from a deceptively simple starting point, eqn 1 can be used to derive the Price equation itself, but also the fundamental equation of population ecology. It therefore defines what eco-evolutionary change is (and is not), mathematically, at its most fundamental. Verbally, it is the statistical interaction between absolute fitness and some characteristic of an individual. We now state this more explicitly (L141).

@@

Figuring this starting point out was not easy; it needed to be done deductively, which inevitably led to many dead ends. Initially, we reasoned that because the Price equation is sufficiently abstract to describe trait change in any population (including groups and higher moments of traits), its recursive properties, and how useful it is for making formal links across disciplines such as physics and information theory (Frank 2015; 2016), we should be able to use it directly to recover population density change. This ended up being a dead end for reasons we will not dwell on (briefly, conservation of total probability -- that entity frequencies need to sum to 1 -- precludes the Price equation from being used to directly to recover absolute population change -- it's not a simple matter of, e.g., setting 'z' to population density). We had different issues when starting from population change and trying to recover the Price equation. There are a lot of ways to make coherent models using these avenues of inquiry, but none that recover universal and exact descriptions of population and evolutionary change. The broader equation, our equation 1, is indeed necessary to derive exact and universal eco-evolutionary change from first principles; it thereby serves as a formal definition for eco-evolutionary change in any system. It perhaps appears simple and obvious in hindsight, as if it could be inferred quickly from existing models by anyone giving it a modicum of thought. But we believe this appearance is highly deceptive; if this formal link were trivial to discover, then someone would have surely pointed it out already. After much investigation, we are confident that no one has, and while there are fundamental equations defining ecological and evolutionary change, separately, our work here is unique in establishing a fundamental equation that defines eco-evolutionary change.

In our revision, we try to clarify the need for our equation more clearly. Our introduction has been completely rewritten in this effort.

@@

Now, in the special case in which z = 1 (which is, again, a unitless count), we recover the total number of individuals, as derived in our manuscript. In this case, Omega is interpreted as individual number (N), and,

Omega = (2 - 1 + 1)(1 + 0) + (1 - 1 + 1)(1 + 0) + (1 - 1 + 1)(1 + 0) = 4

Our population size has increased from 3 to 4 because two individuals produced one offspring each (beta = 1), and one produced two (beta = 2). All of those reproducing individuals died (delta = 1).

We thank Reviewer 2 for raising this query; it gave us pause to reflect and be even more deliberate about what our equation is describing. We now mention in our manuscript why this is not an issue by pointing out that individuals are unitless (L148-L150). We also introduce a new Box 3 (L203-218) walking the reader through a concrete toy example with actual numbers, calculating ecological and evolutionary change, and ecosystem function.

@@

Author response to Reviewer 2 comment 3
======================

This is incorrect; eqn 1 is meaningful as written as a definition of eco-evolutionary change. It is, however, correct that the units associated with 'z' (and therefore 'Omega') are unresolved until 'z' is defined (see our response to Reviewer 2 comment 2 above). But this is no different from the Price equation itself. Our equation doesn't need to recover either the population growth rate equation or the Price Equation to be meaningful. On the contrary, the logic of our manuscript is to develop a unifying framework for eco-evolutionary change. Of course it is good and necessary to prove that we can recover the two most fundamental equations of ecology and evolution. But that is not the end of the story. If our equation only worked through specifying an ecological scenario or an evolutionary scenario, but not both, then it would not fulfil its intended purpose.

We can appreciate why it might feel like our equation 1 works backwards from the fundamental equations of ecology and evolution separately. But this isn't the case, logically or methodologically (see our response to Reviewer 2 comment 1). What we show is that the appropriate starting point for defining eco-evolutionary change is the interaction between births/deaths and individual characteristics. Our starting point isn't arbitrary or contingent; it is exact and universal, and no other equation would work as fundamental (anymore than the Price equation or general population change could be substituted by something else as fundamental equations).

We now emphasise more clearly, in words, the interpretation of equation 1 and how it is meaningful even prior to the derivation of population change or the Price equation (L141-142, L14-16).

@@

Reviewer 2 comment 4 
=========================

Lastly, there are a series of statements made in the Abstract that feel like they either are not followed through on or only weakly mentioned in the Discussion. For example, we are told that this unification “shows how ecological and evolutionary change are reflected in the first and second statistical moments of fitness”, but this is fleshed out in a single sentence in the Discussion (Line 236–238): “Interestingly, the rate of change in ecology and evolution are reflected in the first and second statistical moments of fitness, respectively. Population growth rate reflects mean fitness w ̄ , while the rate of evolutionary change reflects the variance in fitness Var(w).” There appears to be no other mention of this outside the Abstract. Importantly, as the authors note, each of these things are already known – both the equivalence of population growth rate with mean fitness and the rate of evolutionary change being equal to the variance in fitness are found in Fisher (1930).

Author response to Reviewer 2 comment 4 
=========================

We now remove this from the abstract. In the manuscript, we now clarify that our derivation recovers this relationship, which has been pointed out previously (L255-258).

@@

Reviewer 2 comment 5
=========================

Furthermore, the Abstract claims the unification links ecosystem function to population growth rates and evolutionary change, but this is only speculated on in the last paragraph of the Discussion. The authors cite others who have explicitly done this, suggest that their equation might do the same, but there’s no demonstration of it or suggestion of what it might be. Again, this could be because it’s not obvious to me the connection, so the authors need to be a little more explicit in showing these connections.

Author response to Reviewer 2 comment 5
=========================

We are grateful to Reviewer 2 for pointing this out, and we have revised to make our point about ecosystem function clearer with a more concise explanation of our point in our response to Reviewer 2 comment 2. We now elaborate on how Omega can be interpreted as a measure of ecosystem function for a population in a new section (L190-201), and we include an example in a new Box 3 (L203-218).

@@

Reviewer 2 minor comment 2
=========================

Lines 26-28 – I would also cite Betty Smocovitis’ excellent paper here on the unification of biology via evolution: Smocovitis, V. B. (1992). Unifying biology: the evolutionary synthesis and evolutionary biology. Journal of the History of Biology, 25(1), 1-65.

Author response to Reviewer 2 minor comment 2
=========================

This is a good suggestion. We have rewritten the entire introduction, so this sentence no longer exists, but we cite Smocovitis (1992) now in the first sentence of the manuscript (L25).

Reviewer 2 minor comment 3 [ ]
=========================

@@

Lines 45-54 – This historical presentation seems odd to me. My reading of history is that these two fields have always been pretty tightly intertwined; Darwin’s argument relied on Malthusian concepts of population growth, after all, and one of the first concepts of density dependent selection comes from Haldane (1956). But I could be biased. Perhaps a citation after the sentence: “Until recently, population ecology and evolution in particular have taken almost parallel paths for developing theory” to others who have argued this would be helpful.

Author response to Reviewer 2 minor comment 3 
=========================

We have rewritten the introduction, and this sentence no longer exists. In our revision, we acknowledge the long-recognised union between the two disciplines while also pointing out that the rise of eco-evolutionary models is relatively recent following a recognition that ecology and evolution can occur on similar timescales, with citations (L49-52).

@@

After 1960, evolutionary ecology emerged and many ecology textbooks (Pianka 1974, Roughgarden 1979, Begon et al. 1986/2006) included an evolutionary perspective and entire chapters devoted to evolution. However, we think that the pursuit of a true synthesis of evolutionary and ecological change (eco-evo dynamics) has only taken place (roughly) in the last two decades, with several works trying to develop a more unified framework (as far as we know, only one textbook (Hendry 2017) has summarized these eco-evo efforts). Part of the motivation for this synthesis has been the relatively recent appreciation of concurrent ecological and evolutionary dynamics. From Yamamichi, Ellner, and Hairston Jr. (2023), "Although ecology and evolutionary biology have been neighbouring research areas ever since the age of Darwin, the prevailing assumption had long been that ecological processes occur much faster than evolutionary processes".

One of our critiques of eco-evolutionary theory is that there is no formal integration of the foundational axioms shared by both ecology and evolution. Our take is that we should have an eco-evo formalism tout court, which takes a separate but complementary role to existing eco-evolutionary models, which require simplifying assumptions and typically start with ecology or with evolution and then try to include the other one in the model.

Finally, we also agree with Reviewer 2 that classic evolutionary works (like those developed by Haldane) had ecological ingredients. Actually, as we said in our manuscript, our claim is that, contrary to a view among evolutionary biologists (see L259-272), population genetics has ecology at its core, though not in an obvious way.

@@

Reviewer 2 minor comment 4 
=========================

Line 106 – “Fischer” should be spelled “Fisher”.

Author response to Reviewer 2 minor comment 4
=========================

Fixed (Line 95).

Reviewer 2 minor comment 5 [ ]
=========================

@@

Lines 273-274 – If these derivations were included here, it would dramatically clarify my confusion about the usefulness and meaning of Omega in the general formula.

Author response to Reviewer 2 minor comment 5  
=========================

We are reluctant to include too much work still in preparation, or distract from what we believe is the central message of the current manuscript. But we have now included additional derivations in supporting information.

In this new supporting information, we use our eqn 1 to derive (1) density-dependent population growth, (2) multi-level selection (and therefore a version of Hamilton's rule for when altruism is predicted to evolve in a population), and (3) a novel integration of ecology and evolution that demonstrates the relationship between how sociality and fitness (i.e., the extent to which conspecific interactions can muldulate fitness) affect trait change, and how this effect depends on population size -- while simultaneously deriving a classic logistic growth equation.

### 2025-06-18 Supporting_Information_1.Rmd
Edits for resubmission of revision to ecology and evolution

In S5, indices $\beta_{j, i}$, $\delta_{j, i}$, and $z_{j, i}$ identify individual $i$ in group $j$.
For simplicity, we set $w_{j, i} = \beta_{j, i} - \delta_{j, i} + 1$ and let $\Delta z_{j, i} = 0$ (i.e., no transmission bias),

@@

$$\bar{\bar{w}}\Delta = \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right).
\tag{S6}
$$

This recovers the multi-level @Price1972 equation [@Lehtonen2020a] from a starting point of eco-evolutionary change in different groups.
This starting point can be found in @Lehtonen2016 B2.I, which then derives a multi-level selection version of Hamilton's rule predicting the evolution of altruism.

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S11}
$$

We can rewrite the left-hand side of S11,

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i} - w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S12}
$$

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S13}
$$

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right)\bar{w_{i}}. 
\tag{S14}
$$

>>>>

In S5, indices $\beta_{j, i}$, $\delta_{j, i}$, and $z_{j, i}$ identify individual $i$ in group $j$.
We set $w_{j, i} = \beta_{j, i} - \delta_{j, i} + 1$, and for simplicity let $\Delta z_{j, i} = 0$ (i.e., no transmission bias),

@@

$$\bar{\bar{w}}\Delta \bar{\bar{z}} = \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right).
\tag{S6}
$$

This recovers the multi-level @Price1972 equation [@Lehtonen2020a] from a starting point of eco-evolutionary change in different groups.
Equation S6 can be found in @Lehtonen2016 B2.I, who then derives a multi-level selection version of Hamilton's rule predicting the evolution of altruism.

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S12}
$$

We can rewrite the left-hand side of S12,

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i} - w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S13}
$$

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S14}
$$

@@

$$\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right)\bar{w_{i}}. 
\tag{S15}
$$

### 2025-06-18 Supporting_Information_1.tex
Edits for resubmission of revision to ecology and evolution

In S5, indices \(\beta_{j, i}\), \(\delta_{j, i}\), and \(z_{j, i}\)
identify individual \(i\) in group \(j\). For simplicity, we set
\(w_{j, i} = \beta_{j, i} - \delta_{j, i} + 1\) and let
\(\Delta z_{j, i} = 0\) (i.e., no transmission bias),

@@

\[\bar{\bar{w}}\Delta = \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right).
\tag{S6}
\]

This recovers the multi-level Price (\citeproc{ref-Price1972}{1972})
equation (\citeproc{ref-Lehtonen2020a}{Lehtonen 2020}) from a starting
point of eco-evolutionary change in different groups. This starting
point can be found in Lehtonen (\citeproc{ref-Lehtonen2016}{2016}) B2.I,
which then derives a multi-level selection version of Hamilton's rule
predicting the evolution of altruism.

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S11}
\]

We can rewrite the left-hand side of S11,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i} - w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S12}
\]

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S13}
\]

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right)\bar{w_{i}}. 
\tag{S14}
\]

>>>>

In S5, indices \(\beta_{j, i}\), \(\delta_{j, i}\), and \(z_{j, i}\)
identify individual \(i\) in group \(j\). We set
\(w_{j, i} = \beta_{j, i} - \delta_{j, i} + 1\), and for simplicity let
\(\Delta z_{j, i} = 0\) (i.e., no transmission bias),

@@

\[\bar{\bar{w}}\Delta \bar{\bar{z}} = \mathrm{Cov}\left(\bar{w}_{j}, \bar{z}_{j} \right) + \mathrm{E}\left(\mathrm{Cov}_{j}\left(w_{j}, z_{j}\right)\right).
\tag{S6}
\]

This recovers the multi-level Price (\citeproc{ref-Price1972}{1972})
equation (\citeproc{ref-Lehtonen2020a}{Lehtonen 2020}) from a starting
point of eco-evolutionary change in different groups. Equation S6 can be
found in Lehtonen (\citeproc{ref-Lehtonen2016}{2016}) B2.I, who then
derives a multi-level selection version of Hamilton's rule predicting
the evolution of altruism.

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right), z_{i}  \right). 
\tag{S12}
\]

We can rewrite the left-hand side of S12,

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i} - w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S13}
\]

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(w_{i}\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right). 
\tag{S14}
\]

@@

\[\overline{w_{i}\left(1 - \sum_{j=1}^{N}\alpha_{i, j}\right)}\Delta\bar{z} = \mathrm{Cov}\left(w_{i}, z_{i}  \right) - \mathrm{Cov}\left(\sum_{j=1}^{N}\alpha_{i, j}, z_{i}  \right)\bar{w_{i}}. 
\tag{S15}
\]
